---
title: "Using the 'ECweather' package"
author: Braeden Klaver
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ECweather}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE
)

options(rmarkdown.html_vignette.check_title = FALSE)

```

```{r, eval = FALSE}

library(ECweather)

```

While Environment Canada has a tool to download data based on station ID and locations
this is not the best when looking to extract lots of data from different periods of time
over many weather stations, especially if you are looking to automate this process in a 
data pipeline. 

The package ECweather aims to make this process easier by allowing users to extract data
from many time periods, for many weather stations, for hourly, daily, or monthly time frames.

If you are looking to only pull data for one specific station you can also look to the Environment
Canada for one time use, you may find it simpler to manually do this using their tool.

## Required set-up for this package

To use this package you will need to identify the correct station IDs that you are looking for, 
which can be found in the station_inventory dataset included in this package. As well, you 
will need to ensure that the weather station has the data you need. For example, not all weather
stations contain hourly data (where humidity is found) and not all weather stations have been active
throughout the time period this data began to be collected.

```{r, eval = FALSE}

data(ECweather::station_inventory)

```

## Examples of using this package

Once you have identified your required weather station IDs, the time period you are looking for,
and made sure the weather station collects the data you need, you can set up your data call accordingly.

### Pulling daily data

Here we will pull daily data for stations 50 and 52, between 2020 and 2023.

```{r, eval = FALSE}

df <- getECdata(stations = c(50, 52),
                start_year = 2020,
                end_year = 2023,
                timeframe = "daily")

```

### Pulling and saving daily data

In addition to pulling data into the local environment, you can also identify a location to save the output
to use later.

```{r, eval = FALSE}

df <- getECdata(stations = c(50, 52),
                start_year = 2020,
                end_year = 2023,
                timeframe = "daily",
                download = TRUE,
                folder = "path/to/folder")

```

### Pulling hourly data

You can also pull hourly and monthly data.

```{r, eval = FALSE}

df <- getECdata(stations = c(50, 52),
                start_year = 2020,
                end_year = 2023,
                timeframe = "hourly")

```


### Pulling and saving data

In addition to pulling data into the local environment, you can also identify a location to save these files
to use later.

```{r, eval = FALSE}

df <- getECdata(stations = c(50, 52),
                start_year = 2020,
                end_year = 2023,
                timeframe = "daily",
                download = TRUE,
                folder = "path/to/folder")

```

### Extract URL paths

The function 'getECdata' wraps the other function available in the package called 'getECurls'.
If you want to just get these URLs and manage the data load on your own, you can just get a 
list of URLs. Note, this will only generate URLs for one station at a time.

```{r, eval = FALSE}

df <- getECurls(stations = 50,
                start_year = 2020,
                end_year = 2023,
                timeframe = "daily")

```


## Error and warnings messages

### Incorrect timeframe provided

There are only three types of timeframes available for the Weather Canada stations: 
daily, hourly, and monthly. If you do not select one of these options you will get an error
such as:

```
Error in getECurls(id = 1, year_start = 2020, year_end = 2023, timeframe = "annually") : 
  That timeframe is not an option, please select one of: hourly, daily, or monthly.
  
```

### Too many timeframes provided

In order to extract the data the function can only handle one timeframe at a time. If
you provide more than one timeframe in a single pull you will get an error such as:

```
Error in getECurls(id = 1, year_start = 2020, year_end = 2023, timeframe = c("daily",  : 
  Please select one time frame at a time: hourly, daily, or monthly.
```

### Too many station IDs requested

For the function 'getECurls' you can only submit one station ID at a time. If you
request more than one at a time you will get an error such as:

```
Error in getECurls(id = c(1, 3, 4), year_start = 2020, year_end = 2023,  : 
  Please select one weather station ID at a time.
```

### Folder path not provided

For the function 'getECdata' if you are wanting to download the data to a specified folder, but have forgotten
to input a folder you will get an error such as:

```
Error in getECdata(stations = 52, year_start = 2020, year_end = 2023,  : 
  Please provide a folder location to write the data to.
```

### Invalid folder path

For the function 'getECdata' if you are wanting to download the data to a specified folder, but have input
an invalid folder path you will get an error such as:

```
Error in getECdata(stations = 52, year_start = 2020, year_end = 2023,  : 
  Folder path doesn't exist.
```
